## Reproducible Research: Peer Assessment 1

### Loading and preprocessing the data


```{r, echo=TRUE}
activity<-read.csv("activity.csv" , header=TRUE)
head(activity)
```

```{r, echo=TRUE}
dim(activity)
```

```{r, echo=TRUE}
names(activity)
```

```{r, echo=TRUE}
library(lattice)
activity$date<-as.Date(activity$date,format = '%Y-%m-%d')
```

```{r, echo=TRUE}
activity<-read.csv("activity.csv" , header=TRUE)
Total_Steps_By_Day <-aggregate(steps~date, data=activity,sum, na.rm=TRUE)
head(Total_Steps_By_Day)

```


### What is mean total number of steps taken per day?

```{r, echo=TRUE}
barplot(Total_Steps_By_Day$steps, names.arg =Total_Steps_By_Day$date, main="Number of Steps  per Day",xlab="Date", ylab="Steps", col="red")
```

```{r, echo=FALSE}
activity<-read.csv("activity.csv", header=TRUE)
Total_Steps_By_Day <-aggregate(steps~date, data=activity,sum, na.rm=TRUE)
meantotal<-mean(Total_Steps_By_Day$steps)
medianTotal<-median(Total_Steps_By_Day$steps)
```


On average, the number of steps taken per day is `r meantotal` and the median is `r medianTotal` 


### What is the average daily activity pattern?

```{r, echo=TRUE}
activity<-read.csv("activity.csv" , header=TRUE)
Average_Steps_By_Interval <-aggregate(steps~ interval , data=activity,mean, na.rm=TRUE)
plot(Average_Steps_By_Interval$interval, Average_Steps_By_Interval$steps, type="l",xlab="interval",ylab="average steps", main="average number of steps",col="blue")
     
```

```{r, echo=TRUE, results="hide"}
Average_Steps_By_Interval$interval[which.max(Average_Steps_By_Interval$steps)]
```

on average, the 5-minute interval that contains the maximum number of steps is `r Average_Steps_By_Interval$interval[which.max(Average_Steps_By_Interval$steps)]`


### Imputing missing values


```{r, echo=TRUE}

sum(!complete.cases(activity))

```

```{r, echo=TRUE}
activity <- merge(activity,Average_Steps_By_Interval, by = "interval", suffixes = c("",".y"))
nas <- is.na(activity$steps)
activity$steps[nas] <- activity$steps.y[nas]
activity_new <- activity[, c(1:3)]
```

```{r, echo=TRUE}
steps_per_day<-aggregate(steps ~ date, activity_new, sum)
barplot(steps_per_day$steps, names.arg=steps_per_day$date, xlab="date", ylab="steps", main="Number of steps per day", col="red")

```

```{r, echo=FALSE}
mean<-mean(steps_per_day$steps)
median<-median(steps_per_day$steps)
```

On average, the number of steps taken per day is `r mean` and the median is `r median`


### Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
library(lattice)
activity_new$date<-as.Date(activity_new$date,format = '%Y-%m-%d')
activity_new$day <- ifelse(weekdays(activity_new$date) %in%  c("Saturday", "Sunday"),'weekend','weekday')
head(activity_new)
```

```{r, echo=TRUE}
steps_by_interval_and_Daytype<-aggregate(steps~ interval+day,activity_new,FUN="mean") 
head(steps_by_interval_and_Daytype)
```

```{r, echo=TRUE} 
library(lattice)
xyplot(steps ~ interval | day, steps_by_interval_and_Daytype, layout = c(1, 2), type = "l", col="blue", xlab="interval", ylab="average steps", main="average steps pattern by weekday/weekend ")
```

     


